---
title: "Portfoliomanagement and Financial Analysis - Assignment 6"
subtitle: "Submit until Monday 2020-11-02, 13:00"
author: "Schädler, Philipp"
output: html_notebook
---
  
```{r setup}
#remotes::install_github("braverock/FactorAnalytics",  build_vignettes = TRUE, force = TRUE)
pacman::p_load(tidyverse,tidyquant,FFdownload,FactorAnalytics,PerformanceAnalytics)
```

**Please** remember to put your assignment solutions in `rmd` format using **many** chunks and putting readable text in between, similar to my examples given in Research Methods and Assignment 1!

For all exercises: Please use the Assignment-Forum to post your questions, I will try my best to help you along! If you follow the vignettes from `factorAnalytics`, wherever it says `z.score=T`, please exchange it for either `z.score='crossSection'` or `z.score='timeSeries'` depending on the task at hand.


## Exercise 2: Calculating and checking the CAPM cont. (from A05)

As we have seen: the CAPM for small portfolios does not work very well, and so we start using portfolios that get rid of the idiosyncratic risk!
Go to Kenneth French's Homepage  again and download the following datasets: "Portfolios Formed on Market Beta" (where we will use 10 monthly value weighted portfolios formed on beta) and "25 Portfolios Formed on Size and Market Beta" (same thing) as well as the market factor and rf (as before). Now we are going to check the CAPM like famous researchers have done it!
We can use returns as they are in the files (simple returns)!

```{r}

inputlist<-c("F-F_Research_Data_Faktors_CSV.zip","Portfolios_Formed_on_BETA_CSV.zip")
             
#Now process only these files if they can be matched (download only)
FFdownload(output_file = "FFdata.RData", inputlist = inputlist, exclude_daily=TRUE)

load("FFdata.RData")

#Download the Portfolios from Kenneth French's Homepage
portf_mkt_beta <- "https://mba.tuck.dartmouth.edu/pages/faculty/ken.french/ftp/Portfolios_Formed_on_BETA_CSV.zip"
 portf_mkt_beta_csv <- "Portfolios_Formed_on_BETA.csv"
 temp <- tempfile()
download.file(portf_mkt_beta, temp, quiet = TRUE)
portf_mkt_beta <- read_csv(unz(temp, portf_mkt_beta_csv), skip = 15, quote = "\",") %>%
  dplyr::rename(date = "X1") %>%
  mutate_at(vars(-date), as.numeric) %>%
  mutate(date = rollback(ymd(parse_date_time(date, "%Y%m") + months(1))))%>%
  filter(date >= first('1964-01-01') & date <= '2019-12-31')

#Download the market factor and rf (Fama/French 3 Research Factors)
mkt_factors <- "https://mba.tuck.dartmouth.edu/pages/faculty/ken.french/ftp/F-F_Research_Data_Factors_CSV.zip"
 mkt_factors_csv <- "F-F_Research_Data_Factors.CSV"
 temp <- tempfile()
download.file(mkt_factors, temp, quiet = TRUE)
mkt_factors <- read_csv(unz(temp, mkt_factors_csv), skip = 3, quote = "\",") %>%
  dplyr::rename(date = "X1") %>%
  mutate_at(vars(-date), as.numeric) %>%
  mutate(date = rollback(ymd(parse_date_time(date, "%Y%m") + months(1)))) %>%
  filter(date >= first('1964-01-01') & date <= '2019-12-31')


```


a)	Subtract the risk-free rate from the first set of 10 portfolios (only sorted on beta) (Lo 10,., Hi 10) and estimate each stocks beta with the market. Estimate the mean-return for each stock and plot the return/beta-combinations. Create the security market line and include it in the plot! What do you find? (You can split the file in 2-3 different time blocks and see if something changes). * Now we are done with the first-pass regression.*

```{r}

#join data and subtract risk free rate 

ten_portf <- portf_mkt_beta[1:672, -c(2:6)]
ten_portf_joined <- left_join(mkt_factors, ten_portf)

for (i in 6:(ncol(ten_portf_joined))){
ten_portf_joined[,i] <- ten_portf_joined[,i] - ten_portf_joined[,5]
}
ten_portf_rf <- ten_portf_joined[,-(2:5)]

```

```{r}
#Create XTS
mkt_factors_xts <- tk_xts(data = mkt_factors, date_var = date)
ten_portf_rf_xts <- tk_xts(data = ten_portf_rf, date_var = date)

#Calculate Betas for each portfolio
betas_ten_portf <- CAPM.beta(Ra = ten_portf_rf_xts, Rb = mkt_factors_xts[, 1], Rf = 0)

betas_ten_portf

```

Estimate the mean-return for each stock and plot the return/beta-combinations.

```{r}
#Estimate Mean Return
mean_ten_portf_rf_xts <- as.data.frame(lapply(ten_portf_rf_xts, FUN=mean))
mean_ten_portf_rf_xts

#Plot the return/beta-combinations
plot.default(x = betas_ten_portf, xlim=c(0, 2),
             y = mean_ten_portf_rf_xts, ylim=c(0, 1),
             xlab = "Beta", ylab = "Mean Return",
             main = "Return/Beta-combinations")
```

Create the security market line and include it in the plot! What do you find?

```{r}
mean_mkt <- as.data.frame(lapply(mkt_factors_xts[, 1], FUN=mean))
y_mkt <- mean_mkt[1, 1]
plot.default(x = betas_ten_portf, xlim=c(0, 2),
             y = mean_ten_portf_rf_xts, ylim=c(0, 1),
             xlab = "Beta", ylab = "Mean Return",
             main = "Return/Beta-combinations",
             abline(0, y_mkt))
plot.default(x = betas_ten_portf, xlim=c(0, 2), 
             y = mean_ten_portf_rf_xts, ylim=c(0, 10), 
             xlab = "Beta", ylab = "Mean Return",
             main = "Return/Beta-combinations",
             abline(0, y_mkt))

#summary
summary_CAPM_ten_portf <- (table.CAPM(Ra = ten_portf_rf_xts, Rb = mkt_factors_xts[, 1], Rf = 0)[1:9, ])

summary_CAPM_ten_portf

```

(You can split the file in 2-3 different time blocks and see if something changes). * Now we are done with the first-pass regression.*

```{r}


#look for first 10 years
ten_portf_rf_10yrs_xts <- ten_portf_rf[1:120, ] %>%
  tk_xts(date_var = date, silent = TRUE)
betas_ten_portf_rf_10yrs <- CAPM.beta(Ra = ten_portf_rf_10yrs_xts, Rb = mkt_factors_xts[1:120, 1], Rf = 0)
mean_ten_portf_rf_10yrs_xts <- as.data.frame(lapply(ten_portf_rf_10yrs_xts, FUN=mean))
mean_mkt_10yrs <- as.data.frame(lapply(mkt_factors_xts[1:120, 1], FUN=mean))
y_mkt_10yrs <- mean_mkt_10yrs[1, 1]
plot.default(x = betas_ten_portf_rf_10yrs, xlim=c(0, 2),
             y = mean_ten_portf_rf_10yrs_xts, ylim=c(0, 1),
             xlab = "Beta", ylab = "Mean Return",
             main = "Return/Beta-combinations 1964-1974",
             abline(0, y_mkt_10yrs))
summary_CAPM_ten_portf_10yrs <- (table.CAPM(Ra = ten_portf_rf_xts[1:120, ], Rb = mkt_factors_xts[1:120, 1], Rf = 0)[1:9, ])
summary_CAPM_ten_portf_10yrs

```

```{r, echo=FALSE}


#look for 2000-2019
ten_portf_rf_2000_xts <- ten_portf_rf[433:672, ] %>%
  tk_xts(date_var = date, silent = TRUE)
betas_ten_portf_rf_2000 <- CAPM.beta(Ra = ten_portf_rf_2000_xts, Rb = mkt_factors_xts[433:672, 1], Rf = 0)
mean_ten_portf_rf_2000_xts <- lapply(ten_portf_rf_2000_xts, FUN=mean)
mean_ten_portf_rf_2000_xts <- as.data.frame(mean_ten_portf_rf_2000_xts)
mean_mkt_2000 <- lapply(mkt_factors_xts[433:672, 1], FUN=mean)
mean_mkt_2000 <- as.data.frame(mean_mkt_2000)
y_mkt_2000 <- mean_mkt_2000[1, 1]
plot.default(x = betas_ten_portf_rf_2000, xlim=c(0, 2),
             y = mean_ten_portf_rf_2000_xts, ylim=c(0, 1),
             xlab = "Beta", ylab = "Mean Return",
             main = "Return/Beta-combinations 2000-2019",
             abline(0, y_mkt_2000))
summary_CAPM_ten_portf_2000 <- (table.CAPM(Ra = ten_portf_rf_xts[433:672, ], Rb = mkt_factors_xts[433:672, 1], Rf = 0)[1:9, ])
summary_CAPM_ten_portf_2000

plot.default(x = betas_ten_portf, xlim=c(0, 2),
             y = mean_ten_portf_rf_xts, ylim=c(0, 1),
             xlab = "Beta", ylab = "Mean Return",
             main = "Return/Beta-combinations 1964-2019",
             abline(0, y_mkt))
summary_CAPM_ten_portf


```

b)	In the second-pass regression we now regress the average stock returns on the betas estimated before. What do you find in the coefficients and does this contradict the CAPM? Try different time periods again and see what you find. (all of the interpretations are in BKM pp.416f).

```{r}

#Plot return/beta-combination with regression 

plot(x = betas_ten_portf, xlim=c(0, 2),
             y = mean_ten_portf_rf_xts, ylim=c(0, 1),
             xlab = "Beta", ylab = "Mean Return",
             main = "Mean Beta Regression",
             abline(lm(as.numeric(mean_ten_portf_rf_xts) ~ as.numeric(betas_ten_portf))))

lm(as.numeric(mean_ten_portf_rf_xts) ~ as.numeric(betas_ten_portf))

plot(x = betas_ten_portf_rf_10yrs, xlim=c(0, 2),
             y = mean_ten_portf_rf_10yrs_xts, ylim=c(0, 1),
             xlab = "Beta", ylab = "Mean Return",
             main = "Mean Beta Regression 1964-1974",
             abline(lm(as.numeric(mean_ten_portf_rf_10yrs_xts) ~ as.numeric(betas_ten_portf_rf_10yrs))))

lm(as.numeric(mean_ten_portf_rf_10yrs_xts) ~ as.numeric(betas_ten_portf_rf_10yrs))

plot.default(x = betas_ten_portf_rf_2000, xlim=c(0, 2),
             y = mean_ten_portf_rf_2000_xts, ylim=c(0, 1),
             xlab = "Beta", ylab = "Mean Return",
             main = "Mean Beta Regression 2000-2019",
             abline(lm(as.numeric(mean_ten_portf_rf_2000_xts) ~ as.numeric(betas_ten_portf_rf_2000))))



```

```{r}
#There are a number of reasons we expect might the CAPM to fail:
#1. Imperfect measures of the market portfolio
#2. Beta is an incomplete measure of risk
#3. Tax effects
#4. Non - normality of returns
#5. No riskless asset
#6. Divergent borrowing and lending rates
```

c)	Now do the extended second pass regression (regress on betas and residual-sds that you can extract from the regression) and see what you find for different periods. Interpret according to concept check 13.2. One of the (many) problems of the CAPM can be the correlation between residual variances and betas. Calculate and interpret.

```{r}

#Plot return/beta-combination with regression 

res_portf <- residuals(lm(as.numeric(mean_ten_portf_rf_xts) ~ as.numeric(betas_ten_portf)))
plot(x = betas_ten_portf, xlim=c(0, 2),
             y = res_portf, ylim=c(-0.1, 0.1),
             xlab = "Beta", ylab = "Residuals",
             main = "Residuals Beta",
             abline(0,0))


res_portf_10yrs <- residuals(lm(as.numeric(mean_ten_portf_rf_10yrs_xts) ~ as.numeric(betas_ten_portf_rf_10yrs)))
plot(x = betas_ten_portf_rf_10yrs, xlim=c(0.5, 2),
             y = res_portf_10yrs, ylim=c(-0.2, 0.2),
             xlab = "Beta", ylab = "Residuals",
             main = "Residuals Beta 1964-1974",
             abline(0,0))

res_portf_2000 <- residuals(lm(as.numeric(mean_ten_portf_rf_2000_xts) ~ as.numeric(betas_ten_portf_rf_2000)))
plot.default(x = betas_ten_portf_rf_2000, xlim=c(0, 2),
             y = res_portf_2000, ylim=c(-0.5, 0.5),
             xlab = "Beta", ylab = "Residuals",
             main = "Residuals Beta 2000-2019",
             abline(0,0))


```

d)	Try again with 25 portfolios sorted on size and beta. What do you find? Is that interesting? 

```{r}
inputlist1<-c("F-F_Research_Data_Faktors_CSV.zip","25_Portfolios_ME_BETA_5x5_CSV.zip")
             
#Now process only these files if they can be matched (download only)
FFdownload(output_file = "FFdata.RData", inputlist = inputlist1, exclude_daily=TRUE)
load("FFdata.RData")
twentyfive_portf<-(FFdownload$x_25_Portfolios_ME_BETA_5x5$monthly$average_value_weighted_returns)
mkt_factors_xts <- tk_xts(data = mkt_factors, date_var = date)
twentyfive_portf <- twentyfive_portf[-(1:6),]
twentyfive_portf <- twentyfive_portf[-(673:681),]
index(twentyfive_portf) <- index(mkt_factors_xts)

#subtract risk free rate
for (i in 1:(ncol(twentyfive_portf))){
twentyfive_portf[,i] <- twentyfive_portf[,i] - mkt_factors_xts[,4]
}

#Calculate Betas for each portfolio
betas_25_portf <- CAPM.beta(Ra = twentyfive_portf, Rb = mkt_factors_xts[, 1], Rf = 0)
betas_25_portf

#Estimate Mean Return
mean_twentyfive_portf <- as.data.frame(lapply(twentyfive_portf, FUN=mean))
mean_twentyfive_portf

#Plot the return/beta-combinations
plot.default(x = betas_25_portf, xlim=c(0, 2),
             y = mean_twentyfive_portf, ylim=c(0, 1),
             xlab = "Beta", ylab = "Mean Return",
             main = "Return/Beta-combinations (25 Portfolios)")

#Create the security market line and include it in the plot! What do you find?
mean_mkt <- as.data.frame(lapply(mkt_factors_xts[, 1], FUN=mean))
y_mkt <- mean_mkt[1, 1]
plot.default(x = betas_25_portf, xlim=c(0, 2),
             y = mean_twentyfive_portf, ylim=c(0, 1),
             xlab = "Beta", ylab = "Mean Return",
             main = "Return/Beta-combinations",
             abline(0, y_mkt))

#Plot return/beta-combination with regression 
plot(x = betas_25_portf, xlim=c(0, 2),
             y = mean_twentyfive_portf, ylim=c(0, 1),
             xlab = "Beta", ylab = "Mean Return",
             main = "Mean Beta Regression",
             abline(lm(as.numeric(mean_twentyfive_portf) ~ as.numeric(betas_25_portf))))

#Plot return/beta-combination with regression 
res_portf_25 <- residuals(lm(as.numeric(mean_twentyfive_portf) ~ as.numeric(betas_25_portf)))
plot(x = betas_25_portf, xlim=c(0, 2),
             y = res_portf_25, ylim=c(-0.4, 0.4),
             xlab = "Beta", ylab = "Residuals",
             main = "Residuals Beta",
             abline(0,0))


```


  



